### 2015-01-02 doc/thesis/system.tex
Adding text about globals in the typing rules section

\textit{/* \textbf{begin} TODO list */}

\begin{itemize}
\item Refinement with \_ENV
\item Assignment + left-hand side
\item Locals + inference rules + unique and closed table types
\end{itemize}

\textit{/* \textbf{end} TODO list */}

We start with the rule that defines the most unusual feature of
our type system: the refinment of table types.
The rule \textsc{T-REFINE} allow us to add new keys to \emph{open}
and \emph{unique} table types:
\[
\begin{array}{c}
\mylabel{T-REFINE}\\
\dfrac{\begin{array}{c}
       \env_{1}(n) = \{ k_{1}{:}v_{1}, ..., k_{n}{:}v_{n} \}_{open|unique}\\
       \env_{1} \vdash c:t_{1}, \env_{2} \;\;\;
       \not \exists i \in 1..n \; t_{1} \lesssim k_{i} \;\;\;
       \Nil \not\lesssim t_{1}
       \end{array}}
      {\env_{1} \vdash n[c] {<}t{>}:t, \env_{2}[n \mapsto \{ k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}, t_{1}{:}t\}_{open|unique}]}
\end{array}
\]

\textit{/* Should we split into two rules instead of using $open|unique$? */}

@@

We use the refinement of table types to handle the declaration of global variables.
In Lua, the assignment \texttt{v = 5} is syntactic sugar to \texttt{\string_ENV["v"] = 5}
when \texttt{v} is not a local variable, where \texttt{\string_ENV} is a table
that stores the global environment.
For this reason, Typed Lua treats accesses to global variables as field accesses
to an open table in the top-level scope.
For instance,
\begin{center}
\begin{tabular}{l}
$\string_ENV[``x"] \; {<}\String{>} = ``foo";$\\
$\string_ENV[``y"] \; {<}\Integer{>} = 1$\\
\end{tabular}
\end{center}
uses field assignment to add fields $``x"$ and $``y"$ to $\string_ENV$.
We are assuming that $\string_ENV$ is in the environment and has type $\{\}_{open}$.
Therefore, after these field assignments $\string_ENV$ has type
$\{``x":\String, ``y":\Integer\}_{open}$.

We do not allow the refinement of table types to change the type of
a field that is already present in the table type.
For instance,
\begin{center}
\begin{tabular}{l}
$\string_ENV[``x"] \; {<}\String{>} = ``foo";$\\
$\string_ENV[``x"] \; {<}\Integer{>} = 1$\\
\end{tabular}
\end{center}
do not type check, as we are trying to add a field that already exists
in $\string_ENV$.
We could allow the refinement of existing fields, but this could
lead to fields that have a too general type, which could decrease
the ammout of type checking.

\textit{/* We need to close table types before assigning them to fields. */}
\begin{center}
\begin{tabular}{l}
$\string_ENV[``x"] \; {<}\{\}_{unique}{>} = \{\}$
\end{tabular}
\end{center}

@@

This rule is so simple because we left the difficult part to the
type checking of expression lists.
The list of left-hand values is also a kind of expression list.
The difficult part is to check whether there is an expression in
the expression list that messes the environment, and the rule
\textsc{T-LHSLIST} captures this intuition:
\[
\begin{array}{c}
\mylabel{T-LHSLIST}\\
\dfrac{\env \vdash l_{k}:t_{k}, \env_{k} \;\;\;
       merge(\env_{1}, ..., \env_{n}) = \env_{m} \;\;\;
       n = |\;\vec{l}\;|}
      {\env \vdash \vec{l}:t_{1} \times ... \times t_{n} \times \Value{*}, \env_{m}}
\end{array}
\]

@@

We can also use multiple assignments to refine table types:
\begin{center}
\begin{tabular}{ll}
\multicolumn{2}{l}{$\mathbf{local} \; a:\{\}_{unique} = \{\} \; \mathbf{in}$}\\
& \multicolumn{1}{l}{$a[``x"] \; {<}\String{>}, a[``y"] \; {<}\Integer{>} = ``foo", 1$}\\
\end{tabular}
\end{center}

This example type checks because all the environment changes are consistent, and
$``foo" \times 1 \times \Nil{*} \lesssim \String \times \Integer \times \Value{*}$.
The type of the expression list includes the type $\Nil{*}$ in the end
because our type system has typing rules to add this type when
the type of the expression list does not end with a variadic type.
Our type system has this behavior to use $\Nil$ in the place of missing
values, like Lua does.
This example uses the rule \textsc{T-EXPLIST2}, which is really close
to the rule \textsc{T-LHSLIST}:
\[
\begin{array}{c}
\mylabel{T-EXPLIST2}\\
\dfrac{\env \vdash e_{k}:t_{k}, \env_{k} \;\;\;
       merge(\env_{1}, ..., \env_{n}) = \env_{m} \;\;\;
       n = |\;\vec{e}\;|}
      {\env \vdash \vec{e}:t_{1} \times ... \times t_{n} \times \Nil{*}, \env_{m}}
\end{array}
\]

The next example does not type check because it tries to add
the same field to $a$, but with different types:
\begin{center}
\begin{tabular}{ll}
\multicolumn{2}{l}{$\mathbf{local} \; a:\{\}_{unique} = \{\} \; \mathbf{in}$}\\
& \multicolumn{1}{l}{$a[``x"] \; {<}\String{>}, a[``x"] \; {<}\Integer{>} = ``foo", 1$}\\
\end{tabular}
\end{center}

@@

All the examples that we presented until now use the simplest form of
the table constructor: the empty table $\{\}$.
Its inference rule is straightforward:
\[
\begin{array}{c}
\mylabel{T-CONSTRUCTOR1}\\
\env \vdash \{\}:\{\}_{unique}, \env
\end{array}
\]

@@

The inference rule \textsc{T-CONSTRUCTOR2} infers a table type for
the first form:
\[
\begin{array}{c}
\mylabel{T-CONSTRUCTOR2}\\
\dfrac{\env_{i} \vdash p_{i}:k_{i}{:}v_{i}, \env_{i+1} \;\;\;
       n = |\;\vec{p}\;| \;\;\;
       wf(\{k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}\}_{unique}) = t}
      {\env_{1} \vdash \{\;\vec{p}\;\}:t, \env_{n+1}}
\end{array}
\]

@@

The inference rules for expression pairs would be straightforward
if our table types allowed any type in the keys.
The rules \textsc{T-PAIR1}, \textsc{T-PAIR2}, and \textsc{T-PAIR3}
check if the type of the key is a subtype of $\Boolean$, $\Number$,
or $\String$, respectively:
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-PAIR1}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       t_{1} \subtype \Boolean}
      {\env_{1} \vdash [e_{1}] = e_{2}: t_{1}{:}t_{2}, \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-PAIR2}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       t_{1} \subtype \Number}
      {\env_{1} \vdash [e_{1}] = e_{2}: t_{1}{:}t_{2}, \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-PAIR3}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       t_{1} \subtype \String}
      {\env_{1} \vdash [e_{1}] = e_{2}: t_{1}{:}t_{2}, \env_{3}}
\end{array}
\end{array}
\]

The rule \textsc{T-PAIR4} is a fallback for the previous rules,
but it works only if the type of the key does not include $\Nil$:
\[
\begin{array}{c}
\mylabel{T-PAIR4}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \not\subtype t_{1}}
      {\env_{1} \vdash [e_{1}] = e_{2}: \Any{:}t_{2}, \env_{3}}
\end{array}
\]

@@

Besides allowing the refinement of table types through field assignment,
Typed Lua also keeps track whether it is safe or not safe to change the
type of a table.
To do that, our type system always promotes an \emph{open}
table to \emph{closed} and an \emph{unique} table to \emph{open}
when they appear as expressions, or when they enter in a scope that is
not its declaration scope.
The order matters, as going from \emph{unique} to \emph{open}
before going from \emph{open} to \emph{closed} closes \emph{unique} table types.

The rule \textsc{T-IDREAD} uses the predicate \emph{close} to do
this promotion before we use a local variable:
\[
\begin{array}{c}
\mylabel{T-IDREAD}\\
\dfrac{\env_{1}(n) = t}
      {\env_{1} \vdash n_{e}:close(t), \env_{1}}
\end{array}
\]

This rule does not change the type of the local variable because it
just need to ensure that an alias will not get the same type.
For instance, it makes this unsafe example not type check:
\begin{center}
\begin{tabular}{lll}
\multicolumn{3}{l}{$\mathbf{local} \; a:\{``x":\String \cup \Nil\}_{closed} = \{ [``x"] = ``foo" \} \; \mathbf{in}$}\\
& \multicolumn{2}{l}{$\mathbf{local} \; b:\{``x":\String \}_{unique} = \{ [``x"] = ``foo" \} \; \mathbf{in}$}\\
& & \multicolumn{1}{l}{$a = b \;;\; a[``x"] = \mathbf{nil} \;;\; b[``x"] = b[``x"] \;{..}\; ``bar"$}
\end{tabular}
\end{center}

The assignment $a = b$ is not valid because the aliasing changed the
type of $b$ from $\{``x":\String\}_{unique}$ to $\{``x":\String\}_{open}$,
which is not subtype of $\{``x":\String \cup \Nil\}_{closed}$.

@@

Even though the rule \textsc{T-CAST} handles the case that we assign
an alias to a variable, we still have to close this variable due to
safety.
The rule \textsc{T-LOCAL} is one example where we handle this case:
\[
\begin{array}{c}
\mylabel{T-LOCAL}\\
\dfrac{\env_{1} \vdash el:r_{1}, \env_{2} \;\;\;
       r_{1} \lesssim \vec{t} \;\;\;
       closeset(\env_{2}[\vec{n} \mapsto \vec{t}], fav(\mathbf{local} \; \vec{n{:}t} = el \; \mathbf{in} \; s)) \vdash s:\env_{3}}
      {\env_{1} \vdash \mathbf{local} \; \vec{n{:}t} = el \; \mathbf{in} \; s:\env_{3} - \{\vec{n} \mapsto \vec{t}\}}
\end{array}
\]

This rule prevents the following unsafe example to type check:
\begin{center}
\begin{tabular}{lll}
\multicolumn{3}{l}{$\mathbf{local} \; a:\{\}_{unique} = \{\} \; \mathbf{in}$}\\
& \multicolumn{2}{l}{$\mathbf{local} \; b:\{\}_{open} = a \; \mathbf{in}$}\\
& & \multicolumn{1}{l}{$a[``x"] \; {<}\String{>} = ``foo";$}\\
& & \multicolumn{1}{l}{$b[``x"] \; {<}\Integer{>} = 1$}\\
\end{tabular}
\end{center}

@@

Typed Lua uses the refinement of table types to allow programmers to
build modules and objects.
It also allows Typed Lua to type check an object-oriented idiom that
programmers often use it.
\[
\begin{array}{c}
\mylabel{T-METHOD3}\\
\dfrac{\begin{array}{c}
       \env_{1}(n_{1}) = \{k_{i}{:}v_{i}, ..., k_{n}{:}v_{n}\}_{unique|open} \;\;\;
       \env_{1} \vdash n_{2} : l \;\;\;
       \not \exists i \in 1..n \; l \lesssim k_{i}\\
       closeall(\env_{1}[self \mapsto \{k_{i}{:}v_{i}, ..., k_{n}{:}v_{n}\}_{closed}, \vec{n} \mapsto \vec{t}, \ret \mapsto r]) \vdash s:\env_{2}
       \end{array}}
      {\begin{array}{c}
       \env_{1} \vdash \mathbf{fun} \; n_{1}{:}n_{2} \; (\vec{n{:}t}){:}r \; s:\\
       closeset(\env_{1}[n_{1} \mapsto \{k_{i}{:}v_{i}, ..., k_{n}{:}v_{n}, l{:}\Const \; \Self \times \vec{t} \rightarrow r\}_{unique|closed}, fav(\mathbf{fun} \; (\vec{n{:}t}){:}r \; s))
       \end{array}}
\end{array}
\]

@@

Another common idiom that programmers use in Lua is to overload
the input parameter of functions, and use the function \texttt{type}
to execute different actions according to their types.
\[
\begin{array}{c}
\mylabel{T-IF2}\\
\dfrac{\begin{array}{c}
       \env_{1}(n) = t\\
       closeall(\env_{1}[n \mapsto \String]) \vdash s_{1}:\env_{2} \\
       closeall(\env_{1}[n \mapsto filter(t, \String)) \vdash s_{2}:\env_{3}
      \end{array}}
      {\env_{1} \vdash \mathbf{if} \; type(n) == ``string" \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}:closeset(\env_{1}[n \mapsto t], fav(s_{1}) \cup fav(s_{2}))}

@@

\textit{/* How this is valid if projection types do not have subtyping rules? */}

>>>>

We start with the rule that defines the most unusual feature of
our type system: the refinment of table types.
The rule \textsc{T-REFINE} allow us to add new keys to \emph{open}
and \emph{unique} table types:
\[
\begin{array}{c}
\mylabel{T-REFINE}\\
\dfrac{\begin{array}{c}
       \env_{1}(n) = \{ k_{1}{:}v_{1}, ..., k_{n}{:}v_{n} \}_{open|unique}\\
       \env_{1} \vdash c:t_{1}, \env_{2} \;\;\;
       \not \exists i \in 1..n \; t_{1} \lesssim k_{i} \;\;\;
       \Nil \not\lesssim t_{1} \;\;\;
       t_{2} = close(t)
       \end{array}}
      {\env_{1} \vdash n[c] {<}t{>}:t, \env_{2}[n \mapsto \{ k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}, t_{1}{:}t_{2}\}_{open|unique}]}
\end{array}
\]

@@

We use the refinement of table types to handle global variables.
In Lua, the assignment \texttt{v = v + 1} translates to
\texttt{\string_ENV["v"] = \string_ENV["v"] + 1} when \texttt{v}
is not a local variable, where \texttt{\string_ENV} is a table
that stores the global environment.
For this reason, Typed Lua treats accesses to global variables as field accesses
to an open table in the top-level scope.
For instance,
\[
\string_ENV[``x"] \; {<}\String{>} = ``foo" \;;\; \string_ENV[``y"] \; {<}\Integer{>} = 1
\]
uses field assignment to add fields $``x"$ and $``y"$ to $\string_ENV$.
In all the examples we assume that $\string_ENV$ is in the environment and
has type $\{\}_{open}$.
Therefore, after these field assignments $\string_ENV$ has type
$\{``x":\String, ``y":\Integer\}_{open}$.

We do not allow the refinement of table types to change the type of
a field that is already present in the table type.
For instance,
\[
\string_ENV[``x"] \; {<}\String{>} = ``foo" \;;\; \string_ENV[``x"] \; {<}\Integer{>} = 1
\]
do not type check, as we are trying to add a field that already exists
in $\string_ENV$.
We could have used union types to allow the refinement of existing fields,
but this could lead to fields that have a too general type, which could
decrease the ammount of static type checking.
This means that our type system does not allow the table type to change
towards a type that is not a subtype of the previous type.
In this example,
\[
\{``x":\String \cup \Integer\}_{open} \not\subtype \{``x":\String\}_{open}
\]

We also do not allow the refinement of table types to introduce
\emph{open} or \emph{unique} table fields.
For instance,
\begin{center}
\begin{tabular}{l}
$\string_ENV[``x"] \; {<}\{\}_{unique}{>} = \{\}$
\end{tabular}
\end{center}
refines the type of $\string_ENV$ from $\{\}_{open}$ to $\{``x":\{\}_{closed}\}_{open}$,
though we are trying to refine to $\{``x":\{\}_{unique}\}_{open}$.
We made this restriction because we still need to prove that the refinement
of table types is safe.

@@

This rule is so simple because we left the difficult part to the
type checking of expression lists.
The list of left-hand values is also a kind of expression list.
The difficult part is to check whether there is an expression in
the expression list that messes the environment, and the rule
\textsc{T-LHSLIST} captures this intuition:
\[
\begin{array}{c}
\mylabel{T-LHSLIST}\\
\dfrac{\env \vdash l_{k}:t_{k}, \env_{k} \;\;\;
       \env_{m} = merge(\env_{1}, ..., \env_{n}) \;\;\;
       n = |\;\vec{l}\;|}
      {\env \vdash \vec{l}:t_{1} \times ... \times t_{n} \times \Value{*}, \env_{m}}
\end{array}
\]

@@

We can also use multiple assignments to refine table types:
\[
\string_ENV[``x"] \; {<}\String{>}, \string_ENV[``y"] \; {<}\Integer{>} = ``foo", 1
\]

This example type checks because all the environment changes are consistent, and
$``foo" \times 1 \times \Nil{*} \lesssim \String \times \Integer \times \Value{*}$.
The type of the expression list includes the type $\Nil{*}$ in the end
because our type system has typing rules to add this type,
when the type of the expression list does not end with a variadic type.
Our type system has this behavior to use $\Nil$ in the place of missing
values, like Lua does.
This example uses the rule \textsc{T-EXPLIST2}, which is really close
to the rule \textsc{T-LHSLIST}:
\[
\begin{array}{c}
\mylabel{T-EXPLIST2}\\
\dfrac{\env \vdash e_{k}:t_{k}, \env_{k} \;\;\;
       \env_{m} = merge(\env_{1}, ..., \env_{n}) \;\;\;
       n = |\;\vec{e}\;|}
      {\env \vdash \vec{e}:t_{1} \times ... \times t_{n} \times \Nil{*}, \env_{m}}
\end{array}
\]

The next example does not type check because it tries to add
the same field to $\string_ENV$, but with different types:
\[
\string_ENV[``x"] \; {<}\String{>}, \string_ENV[``x"] \; {<}\Integer{>} = ``foo", 1
\]

@@

The simplest form of the table constructor is the empty table $\{\}$.
Its inference rule is straightforward:
\[
\begin{array}{c}
\mylabel{T-CONSTRUCTOR1}\\
\env \vdash \{\}:\{\}_{unique}, \env
\end{array}
\]

@@

The inference rule \textsc{T-CONSTRUCTOR2} infers a table type for
the first form:
\[
\begin{array}{c}
\mylabel{T-CONSTRUCTOR2}\\
\dfrac{\env_{i} \vdash p_{i}:k_{i}{:}v_{i}, \env_{i+1} \;\;\;
       n = |\;\vec{p}\;| \;\;\;
       t = wf(\{k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}\}_{unique})}
      {\env_{1} \vdash \{\;\vec{p}\;\}:t, \env_{n+1}}
\end{array}
\]

@@

The inference rules for expression pairs would be straightforward
if our table types allowed any type in the keys.
The rules \textsc{T-PAIR1}, \textsc{T-PAIR2}, and \textsc{T-PAIR3}
check if the type of the key is a subtype of $\Boolean$, $\Number$,
or $\String$, respectively:
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-PAIR1}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       t_{1} \subtype \Boolean}
      {\env_{1} \vdash [e_{1}] = e_{2}: t_{1}{:}close(t_{2}), \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-PAIR2}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       t_{1} \subtype \Number}
      {\env_{1} \vdash [e_{1}] = e_{2}: t_{1}{:}close(t_{2}), \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-PAIR3}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       t_{1} \subtype \String}
      {\env_{1} \vdash [e_{1}] = e_{2}: t_{1}{:}close(t_{2}), \env_{3}}
\end{array}
\end{array}
\]

The rule \textsc{T-PAIR4} is a fallback for the previous rules,
but it works only if the type of the key does not include $\Nil$:
\[
\begin{array}{c}
\mylabel{T-PAIR4}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \not\subtype t_{1}}
      {\env_{1} \vdash [e_{1}] = e_{2}: \Any{:}close(t_{2}), \env_{3}}
\end{array}
\]

@@

Besides allowing the refinement of table types through field assignment,
Typed Lua also keeps track whether it is safe or not safe to change the
type of a table.
%To do that, our type system always promotes an \emph{open}
%table to \emph{closed} and an \emph{unique} table to \emph{open}
%when they appear as expressions, or when they enter in a scope that is
%not its declaration scope.
%The order matters, as going from \emph{unique} to \emph{open}
%before going from \emph{open} to \emph{closed} closes \emph{unique} table types.
In the last section we mentioned that aliasing \emph{unique} table
types is unsafe.
To avoid unsafe aliasing of \emph{unique} table types, the rule \textsc{T-IDREAD1}
always promotes them to \emph{closed}, and updates the variable in the
environment to \emph{open}:
\[
\begin{array}{c}
\mylabel{T-IDREAD1}\\
\dfrac{\env_{1}(n) = \{k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}\}_{unique}}
      {\env_{1} \vdash n_{e}:\{k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}\}_{closed}, \env_{1}[n \mapsto \{k_{1}{:}v_{1}, ..., k_{n}{:}v_{n}\}_{open}]}
\end{array}
\]

This is the rule that makes the unsafe example from last section to
not type check:
\begin{center}
\begin{tabular}{lll}
\multicolumn{3}{l}{$\mathbf{local} \; a:\{``foo":\Integer\}_{unique} = \{ [``foo"] = 5 \} \; \mathbf{in}$}\\
& \multicolumn{2}{l}{$\mathbf{local} \; b:\{``foo":\Integer \cup \Nil \}_{closed} = a \; \mathbf{in}$}\\
& & \multicolumn{1}{l}{$b[``foo"] = \mathbf{nil}$}
\end{tabular}
\end{center}

This examples does not type check because aliasing $a$ produces the
type $\{``foo":\Integer\}_{closed}$ that is not subtype of
$\{``foo":\Integer \cup \Nil\}_{closed}$, the type of $b$.

@@

We should also keep tracking of \emph{open} locals, as they can
also be unsafe like in the next example:
\begin{center}
\begin{tabular}{lll}
\multicolumn{3}{l}{$\mathbf{local} \; a:\{\}_{unique} = \{\} \; \mathbf{in}$}\\
& \multicolumn{2}{l}{$\mathbf{local} \; b:\{\}_{open} = a \; \mathbf{in}$}\\
& & \multicolumn{1}{l}{$a[``x"] \; {<}\String{>} = ``foo";$}\\
& & \multicolumn{1}{l}{$b[``x"] \; {<}\Integer{>} = 1$}\\
\end{tabular}
\end{center}

@@

Typed Lua uses the refinement of table types to allow programmers to
build modules and objects.
It also allows Typed Lua to type check an object-oriented idiom that
programmers often use it.
\[
\begin{array}{c}
\mylabel{T-METHOD3}\\
\dfrac{\begin{array}{c}
       \env_{1}(n_{1}) = \{k_{i}{:}v_{i}, ..., k_{n}{:}v_{n}\}_{unique|open} \;\;\;
       \env_{1} \vdash n_{2} : l \;\;\;
       \not \exists i \in 1..n \; l \lesssim k_{i}\\
       closeall(\env_{1}[self \mapsto \{k_{i}{:}v_{i}, ..., k_{n}{:}v_{n}\}_{closed}, \vec{n} \mapsto \vec{t}, \ret \mapsto r]) \vdash s:\env_{2}\\
       t_{m} = \{k_{i}{:}v_{i}, ..., k_{n}{:}v_{n}, l{:}\Const \; \Self \times \vec{t} \rightarrow r\}_{unique|open}
       \end{array}}
      {\begin{array}{c}
       \env_{1} \vdash \mathbf{fun} \; n_{1}{:}n_{2} \; (\vec{n{:}t}){:}r \; s:\\
       closeset(\env_{1}[n_{1} \mapsto t_{m}, fav(\mathbf{fun} \; (\vec{n{:}t}){:}r \; s))
       \end{array}}
\end{array}
\]

@@

Another common idiom that programmers use in Lua is to overload
the input parameter of functions, and use the function \texttt{type}
to execute different actions according to their types.
\[
\begin{array}{c}
\mylabel{T-IF2}\\
\dfrac{\begin{array}{c}
       \env_{1}(n) = t\\
       closeall(\env_{1}[n \mapsto \String]) \vdash s_{1}:\env_{2} \\
       closeall(\env_{1}[n \mapsto filter(t, \String)) \vdash s_{2}:\env_{3}\\
       \env_{4} = closeset(\env_{1}[n \mapsto t], fav(s_{1}) \cup fav(s_{2}))
      \end{array}}
      {\env_{1} \vdash \mathbf{if} \; type(n) == ``string" \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}:\env_{4}}

@@
